The Bot on Speaking Terms: The Effects of Conversation Architecture on Perceptions of Conversational Agents

### 2023-02-16 paper.tex
second pass up to 3.3.2

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
%Conversational agents employ natural language to interact with human users. While it is broadly accepted that the specifics of speech interactions affect the perception of communicating parties, 
%Previous research has demonstrated a relationship between the structure of conversational agents' natural language interaction and how these agents are perceived by their users. However, 
%there is a lack of systematic understanding of the connections between the perception of agents and elements of conversation architecture. % are not well understood. 
%Furthermore, there is no clear definition of the aspects of perception that are affected and the elements of conversation architecture. In this paper, we present a synthesized framework demonstrating the connections between perceptions of agents and conversation architecture elements, using a taxonomy of eleven perceptions grouped into four categories, and ten elements grouped into three categories. We highlight the need for further research for the perception of agent's ability and perception of social connection with agent's, especially as it relates to the linguistic style of conversational agents.  Also, there is the need to be consistent in the assessment of agents' perceptions, and consider ethical concerns in design.
Conversational agents are deliberately designed to interact by simulating natural conversation, and the effectiveness of these interactions is shown to be affected by users' perceptions of an agent. While research on human-human and human-agent communication suggests the association between speech specifics and perceptions of communicating parties, there is a lack of systematic understanding of how an agent’s conversation architecture affects users' perceptions. To address this gap, we present a framework that outlines the relationships between elements of agents’ conversation architecture grouped around dialog strategy, content affectiveness, content style and speech format, and agent characteristics grouped around perceptions of interactions with an agent, agent's ability, sociability, and humanness. Synthesized based on literature review (n=57), the framework demonstrates both the identified relationships and the areas lacking empirical evidence. We discuss the implications of the framework for the conversation design and highlight the inconsistency issues with the domain terminology and measurements.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{conversational agents, user perception, conversational architecture, anthropomorphized perceptions, social cues}

@@

%%%% TODO %%%%
% Add headmap of explored relationships - side by side with the connections - show numbers - DONE
% Put the heatmap of impact earlier in the paper - DONE
% Put the tables together - DONE
% show what is explored vs. the connections that had impacts. - DONE
% User perception of agent instead of anthropomorphized perceptions
% Refer to the explored connection in writing, and connections with impacts
% Update to use perception aspects and categories (instead of themes and organizing themes)
% Update introduction
% Include focus areas for future reasearch
% Reviewed recording up to 17:31
% Question about single column vs. double column format - ask Jaisie

@@

\subsubsection*{Perceptions of Agents} To understand the perceptions of agents related to conversation architecture (RQ1), assessments for the perceptions of agents were extracted from each paper. We used inductive thematic analysis to group these assessments into codes based on their meanings. For example, the code likeability of the agent contains assessments such as the questionnaire item "this voice agent was likeable" used by \cite{cuadra2021my}\cmt{[67]}, Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed}'s set of questions on likeability used by \cite{linnemann2018can}\cmt{[15]}, and Subjective Assessment of Speech System Interface (SASSI) questionnaire \cite{hone2000towards}\cmt{sassi}'s set of questions on likeability used by \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}. This step resulted in 83 unique codes. These codes are then organized based similarity of the meanings, resulting in 11 aspects of perceptions. For example, the aspect of \textit{personality traits} of the agent contains measurements such as likeability, friendliness, and warmth. Lastly, the 11 aspects are grouped into 4 categories: \textit{perception of interaction with agent, perception of agent's ability, perception of social connection with agent, and perception of agent's humanness} (Table \ref{tab:perceptions}).

\subsubsection*{Conversation Architecture Elements}
As to the elements of conversation architecture that are relevant to the perception of agents (RQ2), we extracted the conversation architecture used in each paper as codes. Some of the papers used multiple features in their CA design, such as the the anthropomorphic agent used in Seeger et al's study \cite{seeger2021chatbots}\cmt{[35]}. In these composite situations, we broke them down into their individual features. For example, the anthropomorphic agent \cite{seeger2021chatbots}\cmt{[35]} was separated into emotional expressions, is-typing indicator, emoticons, and response delay and captured as codes. 58 unique codes were created as part of this process, capturing features like sentiment-adaptive responses \cite{diederich2019emulating}\cmt{[25]}, lexical alignment \cite{spillner2021talk}\cmt{[18]}, and typos \cite{westerman2019believe}\cmt{[9]}. We then organized these codes into elements based on similarity. For example, the element of \textit{disfluency} contains fillers \cite{jeong2019exploring}\cmt{[10]}\cite{wester2015artificial}\cmt{[14]}, interjections \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}, and repetitions \cite{yang2021effect}\cmt{[72]}. Lastly, these elements are grouped into categories adapted from the social cues taxonomy structure by Feine et al \cite{feine2019taxonomy}. Based on this paper, the Verbal-Content category is mapped to our category of \textit{linguistic content}, and the Verbal-Style category is mapped to our category of \textit{linguistic style}. Given our analysis focuses on the conversational aspects of a CA, the social cues' categories related to embodiment aspects were not used. Also, we collapsed the auditory and invisible categories into our category of \textit{non-linguistic format} (Table \ref{tab:cues}).

\subsubsection*{Relationship Between Perceptions of Agents and Conversation Architecture}
To study the relationship between conversation architecture and conversation architecture, we extracted the connections explored in each paper, using the perception aspects and conversation architecture elements developed in the previous data analysis steps. Each connection's effect was also recorded based on whether an impact was found in the study and the nature of that impact. Out of our review corpus, 265 explored connections between perceptions of agents and conversation architecture were found. To analyze specifically the relationships between the specifics of conversation architecture that affect the perceptions of CAs, 72 connections that did not result in significant relationships were discarded, resulting in 193 relationships for analysis. For example, the connection between the matching style of a CA and user satisfaction was discarded because Hoegen et al. \cite{hoegen2019end}\cmt{[31]} did not find a significant difference between the style matching agent and the non-style matching agent for overall interaction satisfaction. Out of the 193 relationships, 10 were relationships using agents using multiple architecture elements. They were not included in the framework as it is not possible attribute the perceptions of agents to its underlying elements. Overall, 183 relationships between individual architecture elements and perceptions of agents are incorporated into our synthesized framework, visualized as a heatmap to demonstrate their relationships with each other (Figure \ref{fig:heatmap-impact}).

%183 were relationships between individual architecture elements and perceptions. These single architecture element are included in our synthesized frame, 
%CAs using multiple conversation architecture elements were discarded because perceptions are measured at the aggregate level, we cannot attribute the perceptions of agents down to its underlying elements.
%Perceptions impacted by the composite of multiple architecture elements in an agent will be discussed in more details in the Composite Architecture Elements section.

@@

\subsection{Anthropomorphized Perceptions}

As shown in Table \ref{tab:perceptions}, there are four categories defined for the perceptions of conversational agents covering 11 aspects of perception. The details of each aspect are discussed below.

\textbf{Perception of interaction with agent} assesses the overall interaction quality users had with the conversational agent. The three aspects for this perceptions are usability, engagement and satisfaction. \textit{Usability} captures the utilitarian aspects of the interaction, whether it was accurate, easy to use, efficient or helpful. Some commonly used methods to measure usability include the response accuracy portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, or the NASA Task Load Index (NASA-TLX) \cite{hart1988development}\cmt{nasa} to measure cognitive workload. Questions such as "the system is easy to use" or "it is easy to understand the agent" are also used to evaluate usability. \textit{Engagement} on the other hand captures users' emotional reactions to the CA, whether they have enjoyed the conversation, or annoyed or frustrated with the interaction. Some commonly used methods to measure engagement include the annoyance portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, and the Use Engagement Scale (UES) \cite{o2018practical}\cmt{ues}. Questions such as "I enjoyed using the system" or "I felt frustrated with the agent" are also used to evaluate engagement. Lastly, \textit{Satisfaction} captures users' overall satisfaction interacting with the agent. Questions such as "the overall assessment of conversing with the CA was satisfactory" are used to evaluate this perception.

\textbf{Perception of agent's ability} assesses the perceived capabilities of the agent. While there are some elements of system performance in this category, we are mainly concerned with how agents with varied architecture elements but equivalent system capabilities are perceived differently. Specifically, \textit{intelligence} captures the agent's expertise and knowledge, and it is commonly administered through the Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} using the set of questions related to perceived intelligence. Also survey questions such as asking users about the agent's intelligence and domain knowledge are used to evaluate perceived intelligence. \textit{Competence} takes intelligence one step further by examining the agent's ability to put its intelligence into practice. Surveys or qualitative feedback are usually used to assess if the agent is capable and competent, and whether users have confidence in the agent's ability to get the job done. Lastly, \textit{trust} captures agent's truthfulness, credibility and benevolence. Some commonly used methods to evaluate trust include the Trust Propensity Scale \cite{mayer1999effect} and the Individualized Trust Scale (ITS) \cite{wheeless1977measurement}. Questions such as "is the agent honest" and "can I trust the agent with sensitive information" are also used to evaluate trust.

\textbf{Perception of the social connection with agent} assesses the emotional connections that users have with agents. This category includes perception aspects of conversation tone, social presence and intimacy. \textit{Conversation tone} captures the affective impression of the agent's tone, such as empathy and expressiveness. Also, this aspect includes whether the tone used by the agent is persuasive or appropriate for the conversation. \textit{Social presence} captures the sense of connectness and psychological distance users have with an agent. Some aspects in this category include the sense of familiarity or similarity with the agent, and whether users feel the agent behaves like them or have similar attitudes to them. The category of \textit{Intimacy} extends social presence into the realm of the quality of relationships with an agent. Some commonly used methods to assess intimacy include the set of social attraction questions from the interpersonal attraction questionnaire \cite{mccroskey1975development} and the quality of relationship inventory (QRI) \cite{pierce1997assessing}. These questionnaires include questions like "I think the agent could be a friend of mine", or "I feel we could establish a personal relationship with each other".

\textbf{Perception of agent's humanness} assesses the specifics of anthropomorphized perceptions of conversational agents. \textit{Human-likeness} captures whether the agent presented itself as natural and human-like, or artificial and machine-like. The Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} set of questions related to anthropomorphism and the Ascent of Man scale \cite{kteily2015ascent} are commonly used methods to evaluate human-likeness. Survey questions with semantic scales such as "human-like / machine-like" and "artificial / natural" are also commonly used for this aspect of perception. \textit{Personality traits} captures the human characteristics that are attributes to the agent, such as warm, friendly, likeable, and polite. This is commonly collected as qualitative feedback from users, commenting on whether the agent is extroverted or introverted, or the human characteristics they perceive the agent such as funny or witty. Sometimes the Big-5 personality traits questionnaire \cite{gosling2003very} is used to map an agent's disposition on various personality dimensions.

@@

%CW - second pass up to here

As shown in Table \ref{tab:cues}, there are three categories defined for conversation architecture elements across ten different elements. The details of each category and element are discussed below.

The category of \textbf{linguistic content} maps to the content category under verbal cues in Feine et al's taxonomy, which refers to elements expressed with written or spoken words that are related to the strict and literal meaning of a message (i.e. what is said) \cite{feine2019taxonomy}. The theme of \textit{affective language} captures the injection of emotional words or phrases into the agent's utterance, such as affective expressions \cite{seeger2021chatbots}\cmt{[35]}\cite{yang2017perceived}\cmt{[44]}\cite{zhu2022effects}\cmt{[26]}, sentiment-adaptive responses \cite{diederich2019emulating}\cmt{[25]}, and encouraging words \cite{healey2013relating}\cmt{[39]}. \textit{Social talk} captures non-task related conversations with the user to build social connection, such as self-disclosure \cite{lee2020hear}\cmt{[23]} and social dialogue \cite{volkel2021manipulating}\cmt{[68]}\cite{lubold2016effects}\cmt{[86]}. The theme of \textit{humour} is similar to social talk but specific to CA's attempt to include jokes in its utterances. The separation of social talk and humor is based on findings from previous research, where user's attitude towards humour is different between human-human conversations vs. human-agent conversations \cite{clark2019makes}. We would like to capture the relationship between humour and agent perceptions differently from social talk. Lastly, the theme of \textit{agent-initiated content} captures utterances that are driven by the agent without direct prompts from the user. This elements in this theme include conversation repairs \cite{cuadra2021my}\cmt{[67]}\cite{ashktorab2019resilient}\cmt{[88]} and proactive content like eliciting feedback \cite{xiao2021let}\cmt{[73]}.

The category of \textbf{linguistic style} maps to the style category under verbal cues in Feine et al's taxonomy, which refers to elements expressed with written or spoken words that are related to the meaningful deployment of language variation in a message (i.e. how something is said) \cite{feine2019taxonomy}. The theme of \textit{formality} describes the language style used by the conversational agent, whether it is formal vs. casual, such as using honorific expressions to address users \cite{ouchi2019should}\cmt{[59]}. \textit{Alignment} captures the degree that the agent aligns its utterances to the users, such as lexical alignment on the content and structural of the sentences \cite{huiyang2022improving}\cmt{[17]}\cite{linnemann2018can}\cmt{[15]}, as well as agreement with the user \cite{volkel2021examining}\cmt{[69]}. \textit{Elaborateness} captures the sentence complexity and length of the utterances. For example, Roy et al. explored the differences in perception for variations in elaborateness such as "Cloudly, possibility of snow, high: 4, low: -10" vs "Today will be cloudy, with a high of 4 and a low of -10. Snow is predicted" \cite{roy2021users}\cmt{[71]}. Lastly, the category of \textit{disfluency} captures the style of using non-lexical utterances like filler words ("um, uh") \cite{hu2021enhancing}\cmt{[56]}\cite{jeong2019exploring}\cmt{[10]}, or repetition of words within a sentence \cite{yang2021effect}\cmt{[72]}.

Lastly, the category of \textbf{non-linguistic format} maps loosely to the elements in auditory and invisible categories in Feine et al's taxonomy \cite{feine2019taxonomy}. These are auditory or textual variations in the presentation of the agent's utterances. In the theme of \textit{prosodic cue}, it captures vocal qualities like gender of voice \cite{habler2019effects}\cmt{[63]}\cite{jestin2022effects}\cmt{[81]}, and speech rate \cite{choi2020nobody}\cmt{[54]}. \textit{Text variations} include different formats that text-based agents use to present information to users, like emoticons \cite{kim2019comparing}\cmt{[89]}\cite{wilhelm2022keep}\cmt{[28]}, typos \cite{westerman2019believe}\cmt{[9]}, and response delays \cite{gnewuch2022opposing}\cmt{[20]}\cite{seeger2021chatbots}\cmt{[35]}.

Overall the spread of conversation architecture elements relevant to anthropomorphized perceptions are evenly distributed across linguistic content, linguistic style, and non-linguistic format.  Looking at the different themes, there is good coverage of studies on the impact of affective language and prosodic cues on anthropomorphized perceptions. However, it seemed like there is a lack of studies exploring the effect of agent-initiative content on the perception of agents. There are various existing literature on this theme such as conversation repairs \cite{komatani2010online}\cite{reinkemeier2022repair} and agent proactively sharing content with users \cite{dubiel2019inquisitive}\cite{zargham2022understanding}, but they were not included in this systematic literature review because they did not have measures of anthropomorphized perceptions of the CA. This may be because conversations with CAs have historically being driven by the user, with some recent effort to understand the opportunities to be more proactive \cite{cowan2023introduction}, as existing research are focused on the functional aspects of designing an agent instead of exploring into the anthropomorphized perceptions.

@@

& Humour & humorous content, jokes & 5
& \cite{ceha2021can}\cmt{[57]}\cite{go2021conversational}\cmt{[80]}\cite{khooshabeh2011does}\cmt{[37]}\cite{miyamoto2017improving}\cmt{[46]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \Xhline{1.2pt}

@@

& Disfluency & fillers, interjections, repetitions, \newline typos & 7
& \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}\cite{jeong2019exploring}\cmt{[10]}\cite{niewiadomski2013laugh}\cmt{[85]}\cite{wester2015artificial}\cmt{[14]}\cite{westerman2019believe}\cmt{[9]}\cite{yang2021effect}\cmt{[72]}
\\ \cline{2-5}

@@

@@

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
%Conversational agents employ natural language to interact with human users. While it is broadly accepted that the specifics of speech interactions affect the perception of communicating parties, 
%Previous research has demonstrated a relationship between the structure of conversational agents' natural language interaction and how these agents are perceived by their users. However, 
%there is a lack of systematic understanding of the connections between the perception of agents and elements of conversation architecture. % are not well understood. 
%Furthermore, there is no clear definition of the aspects of perception that are affected and the elements of conversation architecture. In this paper, we present a synthesized framework demonstrating the connections between perceptions of agents and conversation architecture elements, using a taxonomy of eleven perceptions grouped into four categories, and ten elements grouped into three categories. We highlight the need for further research for the perception of agent's ability and perception of social connection with agent's, especially as it relates to the linguistic style of conversational agents.  Also, there is the need to be consistent in the assessment of agents' perceptions, and consider ethical concerns in design.
%[AK Edit] Conversational agents are deliberately designed to interact by simulating natural conversation, and the effectiveness of these interactions is shown to be affected by users' perceptions of an agent. While research on human-human and human-agent communication suggests the association between speech specifics and perceptions of communicating parties, there is a lack of systematic understanding of how an agent’s conversation architecture affects users' perceptions. To address this gap, we present a framework that outlines the relationships between elements of agents’ conversation architecture grouped around dialog strategy, content affectiveness, content style and speech format, and agent characteristics grouped around perceptions of interactions with an agent, agent's ability, sociability, and humanness. Synthesized based on literature review (n=57), the framework demonstrates both the identified relationships and the areas lacking empirical evidence. We discuss the implications of the framework for the conversation design and highlight the inconsistency issues with the domain terminology and measurements.
Conversational agents mimic natural conversation to interact with users. Since the effectiveness of interactions strongly depends on users’ perception of agents, it is crucial to design agents’ behaviors to provide the intended user perceptions. Research on human-agent and human-human communication suggests that speech specifics are associated with perceptions of communicating parties, but there is a lack of systematic understanding of how speech specifics of agents affect users' perceptions. To address this gap, we present a framework outlining the relationships between elements of agents’ conversation architecture (dialog strategy, content affectiveness, content style and speech format) and aspects of users’ perception (interaction, ability, sociability and humanness). Synthesized based on literature reviewed from the domains of HCI, NLP and linguistics (n=57), this framework demonstrates both the identified relationships and the areas lacking empirical evidence. We discuss the implications of the framework for conversation design and highlight the inconsistencies with terminology and measurements.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%\begin{CCSXML}
%<ccs2012>
% <concept>
%  <concept_id>10010520.10010553.10010562</concept_id>
%  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
%  <concept_significance>500</concept_significance>
% </concept>
% <concept>
%  <concept_id>10010520.10010575.10010755</concept_id>
%  <concept_desc>Computer systems organization~Redundancy</concept_desc>
%  <concept_significance>300</concept_significance>
% </concept>
% <concept>
%  <concept_id>10010520.10010553.10010554</concept_id>
%  <concept_desc>Computer systems organization~Robotics</concept_desc>
%  <concept_significance>100</concept_significance>
% </concept>
% <concept>
%  <concept_id>10003033.10003083.10003095</concept_id>
%  <concept_desc>Networks~Network reliability</concept_desc>
%  <concept_significance>100</concept_significance>
% </concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

\ccsdesc{TBD}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{conversational agents, user perception, conversational architecture, anthropomorphized perceptions}

@@

@@

\subsubsection*{Perceptions of Agents} To understand the perceptions of agents related to conversation architecture (RQ1), assessments for the perceptions of agents were extracted from each paper. We used inductive thematic analysis to group these assessments into codes based on their meanings. For example, the code likeability of the agent contains the questionnaire item "this voice agent was likeable" used by \cite{cuadra2021my}\cmt{[67]}, Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed}'s set of questions on likeability used by \cite{linnemann2018can}\cmt{[15]}, and Subjective Assessment of Speech System Interface (SASSI) questionnaire \cite{hone2000towards}\cmt{sassi}'s set of questions on likeability used by \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}. This step resulted in 83 unique codes. These codes are then organized based similarity of the meanings, resulting in 11 aspects of perceptions. For example, the aspect of personality traits of the agent contains measurements such as likeability, funniness, and kindness. Lastly, the 11 aspects are grouped into 4 categories: perception of interaction with agent, perception of agent's ability, perception of sociability with agent, and perception of agent's humanness (Table \ref{tab:perceptions}).

\subsubsection*{Conversation Architecture Elements}
As to the elements of conversation architecture that are relevant to the perception of agents (RQ2), we extracted the specifics of conversation architecture used in each paper. Some of the papers used multiple features in their CA design, such as the the anthropomorphic agent used in Seeger et al's study \cite{seeger2021chatbots}\cmt{[35]}. In these composite situations, we broke them down into their individual features. For example, the anthropomorphic agent \cite{seeger2021chatbots}\cmt{[35]} was separated into emotional expressions, is-typing indicator, emoticons, and response delay and captured as codes. 58 unique codes were created as part of this process, capturing features like sentiment-adaptive responses \cite{diederich2019emulating}\cmt{[25]}, lexical alignment \cite{spillner2021talk}\cmt{[18]}, and typos \cite{westerman2019believe}\cmt{[9]}. We then organized these codes into elements based on similarity, resulting in 12 elements of conversation architecture. For example, the element of disfluency contains fillers \cite{jeong2019exploring}\cmt{[10]}\cite{wester2015artificial}\cmt{[14]}, interjections \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}, repetitions \cite{yang2021effect}\cmt{[72]} and typos \cite{westerman2019believe}\cmt{[9]}. These elements are grouped into 4 categories: dialog strategy, content affectiveness, content style, and speech format (Table \ref{tab:cues}).

\subsubsection*{Relationship Between Perceptions of Agents and Conversation Architecture}
To study the effect of conversation architecture on users' perceptions (RQ3), we extracted the connections explored in each paper, using the perception aspects and conversation architecture elements developed in the previous data analysis steps. Each connection's effect was also recorded based on whether an association was found in the study and the nature of that association. Out of our review corpus, 265 connections between perceptions of agents and conversation architecture were explored in literature. To analyze the specifics of conversation architecture that affect the perceptions of CAs, 72 connections that did not result in observed relationships were discarded, resulting in 193 relationships for analysis. For example, the connection between the matching style of a CA and user satisfaction was removed because Hoegen et al. \cite{hoegen2019end}\cmt{[31]} did not find a significant difference between the style matching agent and the non-style matching agent for overall interaction satisfaction. Out of the 193 relationships, 10 were relationships based on agents using multiple architecture elements (e.g. \cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2021manipulating}\cmt{[68]}). They were not included in the final framework as we could not attribute the perceptions of agents to the underlying core elements. Overall, 183 relationships between individual architecture elements and perceptions of agents are incorporated into our synthesized framework, visualized as a heatmap to demonstrate their relationships with each other (Figure \ref{fig:heatmap-impact}).

@@

\subsection{Perceptions of Agents}

As shown in Table \ref{tab:perceptions}, there are four categories defined for the perceptions of conversational agents covering 11 different aspects of perception. The details of each category and aspect are discussed below.

\textbf{Perception of interaction with agent} assesses the overall interaction quality users had with the conversational agent. The three aspects for this perceptions are usability, engagement and satisfaction. \textit{Usability} captures the utilitarian component of the interaction, whether it was accurate, easy to use, efficient or helpful. Some commonly used methods to evaluate usability include the response accuracy portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, or the NASA Task Load Index (NASA-TLX) \cite{hart1988development}\cmt{nasa} for cognitive workload. Questions such as "the system is easy to use" or "it is easy to understand the agent" are also used to evaluate usability. \textit{Engagement} on the other hand captures users' emotional reactions to the CA, whether they have enjoyed the conversation, or annoyed or frustrated with the interaction. Some commonly used methods to evaluate engagement include the annoyance portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, and the Use Engagement Scale (UES) \cite{o2018practical}\cmt{ues}. Questions such as "I enjoyed using the system" or "I felt frustrated with the agent" are also used to evaluate engagement. Lastly, \textit{Satisfaction} captures users' overall satisfaction interacting with the agent. Questions such as "the overall assessment of conversing with the CA was satisfactory" are used to evaluate this perception.

\textbf{Perception of agent's ability} assesses the perceived capabilities of the agent. Unlike system capabilities that affect agent's performance, the perceived ability category captures perceptions of agents with equivalent system capabilities but varied conversation architecture elements. Specifically, \textit{intelligence} captures the agent's expertise and knowledge, and it is commonly administered through the Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} using the set of questions related to perceived intelligence. Also survey questions such as asking users about the agent's intelligence and domain knowledge are used to evaluate perceived intelligence. \textit{Competence} takes intelligence one step further by examining the agent's ability to put its intelligence into practice. Surveys or qualitative feedback are usually used to assess if the agent is capable and competent, and whether users have confidence in the agent's ability to get the job done. Lastly, \textit{credibility} captures agent's truthfulness, benevolence, and user's confidence in CAs. Some commonly used methods to evaluate trust include the Trust Propensity Scale \cite{mayer1999effect} and the Individualized Trust Scale (ITS) \cite{wheeless1977measurement}. Questions such as "is the agent honest" and "can I trust the agent with sensitive information" are also used to evaluate trust.

\textbf{Perception of the sociability with agent} assesses the emotional connections that users have with agents. This category includes perception aspects of conversation tone, social presence and intimacy. \textit{Conversation tone} captures the affective impression of the agent's tone, such as empathy and expressiveness. Also, this aspect includes whether the tone used by the agent was perceived as friendly, polite or warm. \textit{Social presence} captures the sense of connectness and psychological distance users have with an agent. Some aspects in this category include the sense of familiarity or similarity with the agent, and whether users feel the agent behaves like them or have similar attitudes to them. The category of \textit{Intimacy} extends social presence into the realm of the quality of relationships with an agent. Some commonly used methods to assess intimacy include the set of social attraction questions from the interpersonal attraction questionnaire \cite{mccroskey1975development} and the quality of relationship inventory (QRI) \cite{pierce1997assessing}. These questionnaires include questions like "I think the agent could be a friend of mine", or "I feel we could establish a personal relationship with each other".

\textbf{Perception of agent's humanness} assesses the specifics of anthropomorphized perceptions of conversational agents. \textit{Human-likeness} captures whether the agent presented itself as natural and human-like, or artificial and machine-like. The Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} set of questions related to anthropomorphism and the Ascent of Man scale \cite{kteily2015ascent} are commonly used methods to evaluate human-likeness. Survey questions with semantic scales such as "human-like / machine-like" and "artificial / natural" are also commonly used for this aspect of perception. \textit{Personality traits} captures the human characteristics that are attributes to the agent. This is commonly collected as qualitative feedback from users, commenting on whether the agent is extroverted or introverted, or the perceived personality such as likeable, funny or witty. Sometimes the Big-5 personality traits questionnaire \cite{gosling2003very} is used to map an agent's disposition on various personality dimensions.

@@

@@

& Humour & humorous content, jokes & 6
& \cite{ceha2021can}\cmt{[57]}\cite{go2021conversational}\cmt{[80]}\cite{khooshabeh2011does}\cmt{[37]}\cite{miyamoto2017improving}\cmt{[46]}\cite{niewiadomski2013laugh}\cmt{[85]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \Xhline{1.2pt}

@@

& Disfluency & fillers, interjections, repetitions, \newline typos & 6
& \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}\cite{jeong2019exploring}\cmt{[10]}\cite{wester2015artificial}\cmt{[14]}\cite{westerman2019believe}\cmt{[9]}\cite{yang2021effect}\cmt{[72]}
\\ \cline{2-5}

@@

As shown in Table \ref{tab:cues}, there are four categories defined for conversation architecture across 12 different elements. The details of each category and element are discussed below.

The category of \textbf{dialog strategy} refers to the approach used by a conversational agent to engage in a dialogue with a user, which includes the use of social dialogs, agent-initiated content, and adding delays to responses. Specifically, the element of \textit{social dialogue} captures non-task related conversations with the user to build social connection, such as self-disclosure \cite{lee2020hear}\cmt{[23]} and small talk like chitchats \cite{lubold2016effects}\cmt{[86]}\cite{volkel2021manipulating}\cmt{[68]}. The element of \textit{initiative} captures utterances that are initiated by the agent without direct prompts from the user. For example, the agent is taking the initiative to proactively repair the conversation with a user \cite{ashktorab2019resilient}\cmt{[88]}\cite{cuadra2021my}\cmt{[67]}, or actively elicit feedback from users \cite{xiao2021let}\cmt{[73]}. Lastly, \textit{response delay} refers to the tactic of deliberatively delaying for a certain period of time before an agent responds to users \cite{gnewuch2018faster}\cmt{[19]}\cite{gnewuch2022opposing}\cmt{[20]}. It is mainly used in text-based CAs alongside visual displays of typing indicators \cite{gnewuch2018chatbot}\cmt{[21]}.

\textbf{Content affectiveness} refers to the conversational agent's use of language to convey emotions or to elicit emotions from users. The architecture element of \textit{affective language} captures the injection of emotional words or phrases into the agent's utterance, such as affective expressions \cite{seeger2021chatbots}\cmt{[35]}\cite{yang2017perceived}\cmt{[44]}\cite{zhu2022effects}\cmt{[26]}, sentiment-adaptive responses \cite{diederich2019emulating}\cmt{[25]}, and encouraging words \cite{healey2013relating}\cmt{[39]}. The other element in this category, \textit{humour}, captures an agent's attempt to include jokes in its dialog. Several studies in our reviewed corpus have explored the effect of humour on various perceptions of agents (e.g. \cite{ceha2021can}\cmt{[57]}\cite{khooshabeh2011does}\cmt{[37]}).

The category of \textbf{content style} refers to the variations of language used in a message aside from the content meaning of the message (i.e. how something is said). The element of \textit{formality} describes the linguistic style used by the conversational agent, whether it is formal vs. casual, such as using honorific expressions to address users \cite{ouchi2019should}\cmt{[59]}. \textit{Alignment} captures the degree that the agent aligns its utterances to the users, such as lexical alignment on the content and structural of the sentences \cite{huiyang2022improving}\cmt{[17]}\cite{linnemann2018can}\cmt{[15]}, as well as referring and agreeing with users' utterances \cite{volkel2021examining}\cmt{[69]}. Lastly, the element of \textit{elaborateness} captures the sentence complexity and length of agents' utterances. For example, Roy et al. explored the differences in perception for variations in elaborateness such as "Cloudly, possibility of snow, high: 4, low: -10" vs "Today will be cloudy, with a high of 4 and a low of -10. Snow is predicted" \cite{roy2021users}\cmt{[71]}.

Lastly, the category of \textbf{speech format} captures the non-verbal component of a conversational agent's utterances for both text and voice modalities. \textit{Disfluency} captures 
the use of non-lexical utterances like filler words ("um, uh") \cite{hu2021enhancing}\cmt{[56]}\cite{jeong2019exploring}\cmt{[10]}, or repetition of words within a sentence \cite{yang2021effect}\cmt{[72]}. It also includes the format of using typos \cite{westerman2019believe}\cmt{[9]} for text-based agents. For voice-based agents specifically, the element of \textit{prosody} encompasses vocal qualities like pitch \cite{habler2019effects}\cmt{[63]}\cite{jestin2022effects}\cmt{[81]}, speech rate \cite{choi2020nobody}\cmt{[54]}, and spoken accent \cite{feijoo2021effects}\cmt{[70]}. For text-based CAs, the \textit{text formatting} element includes different formats agents uses to present information to users, such as using capitalization \cite{westerman2019believe}\cmt{[9]} or emoticons \cite{kim2019comparing}\cmt{[89]}\cite{wilhelm2022keep}\cmt{[28]}.

Overall there are similar number of conversation architecture elements explored across the four categories of dialog strategy, content affectiveness, content style, and speech format. As for explorations related to different modalities of agents, there is a noticeable difference in the category of speech format, where architecture elements related to voice-based agents (n=15) are explored more than text-based agents (n=7). Looking across the 12 different elements, there is good coverage of studies on the impact of affective language and prosody cues on the perceptions of agents. However, there are less studies exploring the effect of agent initiated content and response delay on the perception of agents. Specifically for the element of initiative under dialog strategy category, there is literature on conversation repairs \cite{komatani2010online}\cite{reinkemeier2022repair} and agent proactively sharing content with users \cite{dubiel2019inquisitive}\cite{zargham2022understanding}, but many of them did not not include assessments of perceptions of the CA. This may be due to the fact that the usage of agent-initiated content is relatively new, as CA interactions have historically been driven by users. As such, current research is focused on the functional aspects of agents initiating contents instead of exploring the perceptions of agents. As for response delay, there are indications that this element is under-explored in literature. Only 4 studies in our reviewed corpus used an agent with response delay, and the majority of these publications were written by the same authors.

@@

%For humour: Clark et al. \cite{clark2019makes} found that humour is an important conversational characteristic for human-human interactions, but viewed as a novelty feature for human-agent conversations.
